---
title: "HR Data Visualization Q1"
author: "Stephanie Cheng"
output: html_document
date: "2025-01-05"
---

## Load Library
```{r library setup, include=FALSE}
library(lattice)
library(ggplot2)
library(dplyr)
```

## Load Dataset
```{r data setup, include=FALSE}
hr_data <- read.csv("HRDataset_v14.csv")
```

## Q1

The relationship between `ManagerName` and `PerformanceScore`

```{r relationship intro, echo=FALSE}
#See how many employees under each Manager
print(table(hr_data$ManagerName))

```

```{r relationship Plot, echo=FALSE}
hr_data$ManagerFactor <- factor(hr_data$ManagerName)
y_axis <- unique(hr_data$PerfScoreID)

xyplot(PerfScoreID ~ ManagerID | ManagerFactor, 
       data = hr_data, 
       pch = 19, 
       as.table = TRUE,
       main = "Employee Performance Scores by Manager",
       xlab = "",
       ylab = "Performance Score ID",
       strip = strip.custom(par.strip.text = list(cex = 0.6)),
       scales = list(y = list(at = y_axis))
       )
```

## Q2
The overall diversity profile

(1)Gender
```{r diversity_gender, echo=FALSE}
gender_count <- table(hr_data$Sex)
barchart(gender_count,
         main = "Gender Distribution",
         xlab = "Count",
         ylab = "Gender",
         col = c("violet","royalblue"),
         auto.key = list(columns = 2))
```

(2)Ethnicity (Race)
```{r diversity_race, echo=FALSE}
race_data <- hr_data %>%
  group_by(RaceDesc) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Proportion = Count / sum(Count))

ggplot(race_data, aes(x = reorder(RaceDesc, -Count), y = Count, fill = RaceDesc)) +
  geom_bar(stat = "identity") +
  labs(title = "Race Distribution",
       x = "Ethnicity",
       y = "Count",
       fill = "Race") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(race_data)
```

(3)Marital Status
```{r diversity_marital status, echo=FALSE}
marital_data <- hr_data %>%
  group_by(MaritalDesc) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Proportion = Count / sum(Count))

ggplot(marital_data, aes(x = "", y = Count, fill = MaritalDesc)) +
  geom_bar(stat = "identity", width = 1) +  
  coord_polar("y") + 
  labs(title = "Employee Distribution by Marital Status",
       fill = "Marital Status") +
  theme_minimal()

print(marital_data)
```

(4)Citizen
```{r diversity_citizen, echo=FALSE}
citizen_state_data <- hr_data %>%
  group_by(State, CitizenDesc) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Proportion = Count / sum(Count),
         Statelab = as.numeric(factor(State)),
         Citizenlab = as.numeric(factor(CitizenDesc)))

ggplot(citizen_state_data, aes(x = factor(State), y = CitizenDesc, fill = Count)) +
  geom_tile() +
  scale_fill_gradient(low = "springgreen", high = "turquoise4") +
  labs(title = "Employees by State and Citizenship",
       x = "State",
       y = "Citizenship",
       fill = "Employee Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5))


print(citizen_state_data)
```

## Q3
Show the relationship between diverse/non-diverse hires and each recruitment source.

Since there are lots of recruitment source, i will randomly assigns them into 3 different
groups. So that it is readable when plotting the relationship.
```{r recruitment sources, echo=FALSE}
unique_sources <- unique(hr_data$RecruitmentSource)

shuffled_sources <- sample(unique_sources)

# Split into 3 equal groups
num_groups <- 3
split_sources <- split(unique_sources, rep(1:num_groups, each = length(unique_sources) / num_groups))

source_group_map <- setNames(rep(c("Group 1", "Group 2", "Group 3"), each = length(unique_sources) / num_groups),
                             unlist(split_sources, use.names = FALSE))

hr_data$SourceGroup <- source_group_map[hr_data$RecruitmentSource]
hr_data$RecruitmentSource <- factor(hr_data$RecruitmentSource)

```

Plot the relationship 
```{r recruitment sources, echo=FALSE}
COPLOT <- function(group_name) {
  group_data <- hr_data[hr_data$SourceGroup == group_name, ]
  group_data$RecruitmentSource <- factor(group_data$RecruitmentSource)
  
  coplot(FromDiversityJobFairID ~ as.numeric(RecruitmentSource) | RecruitmentSource, 
       data = group_data, 
       panel = function(x, y, ...) {
         panel.smooth(x, y, ...)
         axis(2, at = c(0, 1), labels = c("0", "1"))
       },
       xlab = "Recruitment Source",
       ylab = "Diversity Hire",
       main = "Diversity Hiring by Recruitment Source")
}
COPLOT("Group 1")
COPLOT("Group 2")
COPLOT("Group 3")
```

## Q4

Check who is going to terminate (Outline Source: ChatGPT)

```{r prediction terminate, echo=FALSE}
hr_data$Termd <- as.factor(hr_data$Termd)

selected_data <- hr_data %>%
  select(Employee_Name, EmpID, Termd, PerfScoreID, EngagementSurvey, DaysLateLast30, Absences)

scatter_matrix <- ggpairs(
  selected_data,
  columns = c("PerfScoreID", "EngagementSurvey", "DaysLateLast30", "Absences"),
  aes(color = Termd, alpha = 0.7),
  lower = list(continuous = "points"),
  upper = list(continuous = "cor"),
  title = "Termination Status"
)

print(scatter_matrix)
```

## Q5

Check the pay equality by factors 
```{r pay, echo=FALSE}
pay_data <- hr_data %>%
  select(Department, Position, GenderID, Salary)

print(pay_data)
#by Department
ggplot(pay_data, aes(x = Department, y = Salary, fill = Department)) +
  geom_boxplot(outlier.color = "red", outlier.size = 2, alpha = 0.9) +
  labs(
    title = "Salary Distribution by Department",
    x = "Department",
    y = "Salary"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#by Position
ggplot(pay_data, aes(y = Salary, fill = Position)) +
  geom_boxplot(outlier.color = "red", outlier.size = 2, alpha = 0.9) +
  labs(
    title = "Salary Distribution by Position",
    x = "Position",
    y = "Salary"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

#by Gender
ggplot(pay_data, aes(x = as.factor(GenderID), y = Salary, fill = as.factor(GenderID))) +
  geom_boxplot(outlier.color = "red", outlier.size = 2, alpha = 0.9) +
  labs(
    title = "Salary Distribution by Gender",
    x = "Gender",
    y = "Salary"
  ) +
  theme_minimal()

```
